---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(data.table)
DT = fread("activity.csv")[,date:=as.Date(date)]
```
## What is the mean and total number of steps taken per day?
### Histogram of total steps for each day
```{r}
library(ggplot2)
sum_stats = na.omit(DT[,list(sum=sum(steps)), by=date])
ggplot(sum_stats, aes(date, sum)) + geom_bar(stat="identity")
```

### Table of mean and median steps for each day
```{r}
center_stats = DT[,list(mean=mean(steps), median=as.double(median(steps))), by=date]
print(center_stats)
```

## What is the average daily activity pattern?
### Time Series Plot
```{r}
center_stats = DT[,list(mean=mean(steps, na.rm=TRUE)), by=interval]
ggplot(center_stats, aes(interval, mean)) + geom_line()
```

### Time interval with maximum mean number of steps
```{r}
center_stats_max = center_stats[mean==max(mean)]
with(center_stats_max, print(sprintf("max mean is %f at 5-minute interval %g", mean, interval)))
```

## Imputing missing values

### The number of rows with NA:
```{r}
nrow(DT[is.na(steps)])
```

### The strategy for filling out the missing number of steps for an interval on a given day will be to take the mean value for that interval across all days.
```{r, results='hide'}
DT[,c("steps_raw", "steps") := list(steps, as.double(steps))][is.na(steps), steps := center_stats$mean]
```

### Histogram of total steps for each day after missing values were imputed
```{r}
sum_stats = na.omit(DT[,list(sum=sum(steps)), by=date])
ggplot(sum_stats, aes(date, sum)) + geom_bar(stat="identity")
```

### Print table of median and mean for each day before and after imputing

We inserted the mean value across all days as the imputed value for a given interval. This means that for days that already had values, their mean and median hasn't changed, as illustrated by all the zeroes in the `mean_diff` and `median_diff` columns.


```{r}
options(width=119)
center_stats = DT[,list(mean=mean(steps), median=as.double(median(steps)), mean_raw=mean(steps_raw), median_raw = as.double(median(steps_raw))), by=date][,c("mean_diff", "median_diff") := list(mean - mean_raw, median - median_raw)]
print(center_stats)
```

## Are there differences in activity patterns between weekdays and weekends? YES!

### Create a new weekday/weekend factor variable
```{r, results='hide'}
weekend = c("Saturday", "Sunday")
levels = c(FALSE,TRUE)
labels = c("weekday", "weekend")
DT[,w:=factor(weekdays(date) %in% weekend, levels, labels)]
```

### Graph weekday vs. weekend
```{r, results='hide'}
center_stats = DT[,list(mean=mean(steps, na.rm=TRUE)), by=c("interval","w")]
ggplot(center_stats, aes(interval, mean)) + geom_line() + facet_grid(w ~ .)
```